Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-07-05 analysis/paper/paper.Rmd
updates for introduction and references

Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005]. Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@Wang2007]. Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age. We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale. Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005]. Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Given2004; @Mullins2011; @Silliman2001; @Rubertone2000; @Torrence2000; @Torrence2000a; @Voss2005].

Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced. Direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local communities remains unanswered. Indirect effects of colonialism is addressed recently to discuss the impact on the local indigenous societies in the periphery of colonial control [@Trabert2017]. Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012]. Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period. Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century. Northeastern Taiwan can provide a good example to explore the indirect impacts on indigenous societies caused by different foreign groups in different periods by taking a long-term perspective research.

@@

@@

Indirect influence of colonialism has been proposed recently and some studies reveal how peripheral areas had been influenced by the colonial activities or involved in the colonial economy centered in the major European colonies [@Trabert2017]. Kiwulan in Northeastern Taiwan is an especially important case study where the location was relatively isolated and peripheral that can provide valuable insight into the discussion of influence of indirect colonial contacts. Foreign trade ornaments were viewed as prestige goods and played an important role in the cultural context of indigenous societies in northeastern Taiwan according to the use context as grave goods and historical documents written by foreign record keeper. As a status marker, the frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods. Greater amount and diversity ornament types during the European contact period reflects the international stimulation of exchange in a colonial context. Those ornaments may convey different meaning with exotic and colonizer image that increases more competitions between aggrandized individuals for prestige and wealth accumulation, which might result in a increasing social inequality. The spatial pattern of ornaments at Kiwulan during European contact presents multiple clusters that is significantly different from a random distribution pattern. This study demonstrate that foreign trade ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations and give insights into the emergence of social inequality stimulated by the European colonization. It also shows the agency of indigenous people to incorporate and recontextualize the ornaments into their culture system. During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwalan population. This is a case study that focuses on trade ornaments and more studies can be done to have a comparison with other archaeological evidence. Further studies to explore the extent of social inequality and the shift in social organization could be conducted by examining multiple line of evidence from Kiwulan, including pottery production and mortuary practices.

>>>>

1. background of the topic

Taiwan island became part of a global trading network during the early 17th century because of western European expansion that connected European and Asia. Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced. Direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local indigenous communities remains unanswered. Indirect effects of colonialism is emphasized increasingly to discuss the colonial impact on the indigenous people in the periphery of colonial control [@TorrenceandClarke2000; @Trabert2017]. In many parts of the world, foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts might cause significant transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997; @Junker1993; @Mitchell2000; @Silliman2005]. Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed to emphasize the ability of active adoptions of foreign effects, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Dietler2015; @Given2004; @Rubertone2000; @Silliman2001; @Torrence2000; @TorrenceandClarke2000; @Voss2005]. Studying the consumption of foreign goods is one way to understand indigenous experience of the colonial encounter and explore the agency of indigenous people by examining how those goods were distributed and used when obtained [@Dietler2005; @Mullins2011; @Scaramelli2005].

Taiwan was colonized by the Spanish and the Dutch subsequently from 1624 to 1662, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012]. However, the sea provides a way through which indigenous people could interact with other groups from different regions, including other native indigenous groups, foreign traders such as the Han people of China, and Europeans. Large-scale and frequent trade activities brought a wide variety of trade objects into indigenous communities in Yilan through local regional exchange network. Among those trade goods, ornament such as glass and stone beads was one of the common foreign goods found at the settlement sites during this period [@Chen2007; @Li2014; @NMTH2005; @Wang2007]. Body adornment is usually viewed as a signal in indigenous social life to represent and reinforce their status through the body as the scene to display [@Joyce2005; @Scaramelli2005]. The consumption of trade beads in Southeast Asia during the Iron Age were usually associated with increasing social stratification or socio-political complexity that influenced [@Bellina2014; @Carter2016; @Francis2002; @Theunissen2000].

identify the value of the beads

@Theunissen2000 <---This has never been explored in any depth despite the fact it has been recognized for some time that the introduction of new prestige goods from external sources may cause restructuring of existing prestige goods exchange and subsequent increases in social complexity (Higham 1984: 250-4). --->

@Mullins2011 <---suggest that beads were one class of goods that continued and embellished indigenous value systems that stressed the powers embedded in shiny objects such as quartz crystals, and the European use of beads in items such as rosaries may have reinforced their powers in the eyes of indigenous people.--->

@Carter2016 <---has also proposed that these lower quality beads may instead be related to intensified trade with South Asia and the importation of mass-produced beads from this location. --->

My research will explore the evidence at Kiwulan by addressing the question of whether European contact and influence on local trade networks stimulated an increase in social inequality among the inhabitants of
Kiwulan. In this project, I intend to test the hypothesis that European contact in northeastern Taiwan stimulated a change in indigenous social organization at Kiwulan.

2. debate or problem have not solved yet
3. how this case can answer it
4. what are you going to do

@@

Before European arrival, the native people had already established their own trading networks but the arrival of Europeans and their trade goods in the 17th century built on, but ...

Foreign ornaments including glass beads, stone beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@Wang2007]. Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan during the transition from the Neolithic to the Iron Age. We observed a greater quantity and variety of foreign goods in the early 17th century during the European colonization period that involved the whole island in an international trade network on a large scale.

Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period. Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century. Northeastern Taiwan can provide a good example to explore the indirect impacts on indigenous societies caused by different foreign groups in different periods by taking a long-term perspective research.

@Mullins2011---<They examine the ways gift-giving worked in the Middle Orinoco region of Venezuela, where missionaries hoped to secure indigenous alliances by providing goods for which they perceived a native demand. One class of goods missionaries provided was alcohol, which was used commonly in feasts and ritual contexts, but interjecting externally produced alcohol unseated existing production and consumption systems. Prior to colonization, drink had been produced by indigenous domestic units to enhance a family's prestige, but the introduction of imported spirits compelled hosts to obtain them through exchange, which forced increasing reliance on the production of cash crops and increased debt among indigenous consumers. This shift rendered drink “a commodity—a product that had to be bought, rather than produced,” so even though European alcohol was consumed within existing consumer contexts it led to dependency and exploitation--->

@@

Indirect influence of colonialism has been proposed recently and some studies reveal how peripheral areas had been influenced by the colonial activities or involved in the colonial economy centered in the major European colonies [@Trabert2017]. Kiwulan in Northeastern Taiwan is an especially important case study where the location was relatively isolated and peripheral that can provide valuable insight into the discussion of influence of indirect colonial contacts. Foreign trade ornaments were viewed as prestige goods and played an important role in the cultural context of indigenous societies in northeastern Taiwan according to the use context as grave goods and historical documents written by foreign record keeper. As a marker for status, the frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different culture contact periods. Greater amount and diversity ornament types during the European contact period reflects the international stimulation of exchange in a colonial context. Those ornaments may convey different meaning with exotic and colonizer image that increases more competitions between aggrandized individuals for prestige and wealth accumulation, which might result in a increasing social inequality. The spatial pattern of ornaments at Kiwulan during European contact presents multiple clusters that is significantly different from a random distribution pattern. This study demonstrate that foreign trade ornaments could be a proxy to detect the indirect colonial influence on local indigenous populations and give insights into the emergence of social inequality stimulated by the European colonization. It also shows the agency of indigenous people to incorporate and recontextualize the ornaments into their culture system. During the Chinese contact period, the decreasing frequency and diversity of ornaments showing a decline in the production, use and discard of ornaments at Kiwulan, which might be related to the smaller scale of exchange network with limited sources of ornaments or the overall decline of Kiwalan population. This is a case study that focuses on trade ornaments and more studies can be done to have a comparison with other archaeological evidence. Further studies to explore the extent of social inequality and the shift in social organization could be conducted by examining multiple line of evidence from Kiwulan, including pottery production and mortuary practices.
